A comparison of design-based and model-based approaches for finite population spatial data.

### 2021-11-12 inst/manuscript/manuscript.Rmd
section 3 updates

Performance of the four sampling-analysis combinations were evaluated in 36 different simulation scenarios. The 36 scenarios resulted from the crossing of three sample sizes, two location layouts, two response types, and three proportions of dependent random error. The three sample sizes ($n$) were $n = 50, n = 100,$ and $n = 200$. Samples were always selected from a population size ($N$) of $N = 900$. The two location layouts were random and gridded. Locations in the random layout were selected randomly from the unit square ($[0, 1] \times [0, 1]$). Locations in the gridded layout were selected randomly on a fixed grid from the unit square. The two response types were normal and lognormal. For the normal response type, the response was simulated using mean-zero random errors with the exponential covariance (Equation$~$\ref{equation:expcov}) for varying proportions of dependent random error. The proportion of dependent random error is represented by $\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)$, where $\sigma^2_1$ and $\sigma^2_2$ are from Equation$~$\ref{equation:expcov}. The total variance ($\sigma^2_1 + \sigma^2_2$) was always 2. The range was always $\sqrt{2} / 3$, which means that the correlation in the dependent random error decays to nearly zero at the largest possible distance between two units in the domain. For the lognormal response type, the response was first simulated using the same approach as for the normal response type, except that the total variance was 0.6931 instead of 2. The response was then exponentiated, yielding a total variance of 2 on the exponentiated scale. The lognormal responses were used to evaluate performance of the sampling-analysis approaches for data that were skewed.

a crossed design with the simulation parameters given in Table \ref{tab:parmtab} for a total of 36 scenarios. All scenarios used exponential correlation with a  $\sqrt{4} / 3$ for $N = 900$ response values simulated on the unit square in either random locations (Layout = Random) or gridded locations (Layout = Gridded). The mean for the spatial process generating the response was set to zero.

<!-- The range was chosen so that the decayed to nearly zero at the largest distance possible in the domain, $\sqrt{2}$, otherwise known as the "effective" range (for the exponential covariance, the effective range is $3\phi$). -->

For the lognormal scenarios, the response values were simulated using the specified correlation parameters using a normal distribution and were subsequently exponentiated. A total variance of 0.6931 and a mean of 0 on the normal scale yielded a total variance of 2 and a mean of 1.414 after exponentiation. Therefore, when the model-based methods were used for lognormal response, the correlation was mis-specified. We chose to simulate values with a lognormal distribution so that we could test the model-based analysis approach with a mis-specified model and so that we could test both analysis approaches on data that exhibits a large amount of skewness.

```{r, echo = FALSE, results = "asis"}
n <- c("50", "100", "200")
site_locations <- c("Random", "Gridded", "")
psill_ratio <- c("0", "0.5", "0.9")
resp_type <- c("Normal", "Lognormal", "")
parm_tab <- rbind(n, site_locations, psill_ratio, resp_type)
parm_tab[c(2, 4), 3] <- "-"
row.names(parm_tab) <- c("Sample Size (n)", "Location Layout", 
                         "Proportion of Dependent Error", "Response Type")
# parm_tab <- parm_tab[c(1, 2, 4, 3), ]
parm_xtab <- xtable(parm_tab, caption = "\\label{tab:parmtab} Simulation scenario options. All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios. In each simualtion scenario, the mean was zero and the variance was two.")
align(parm_xtab) <- "r|lll"
print(parm_xtab, include.colnames = FALSE, hline.after = c(0, nrow(parm_xtab)), comment = FALSE)
```

```{r, figlog, out.width = "100%", fig.cap = "Lognormal distribution with a mean of 0 on the log scale and variance of 2 on the log scale. The distribution exhibits moderate to strong right-skewness.", fig.keep = "none"}
mu <- 0; var <- 2
xvar <- seq(0.001, 10, by = 0.01)
yvar <- dlnorm(xvar, meanlog = 0, sdlog = sqrt(var))
df <- tibble(xvar = xvar, yvar = yvar)
ggplot(df, aes(x = xvar, y = yvar)) +
  geom_line()
```

There were 2000 simulation trials for each of the 36 parameter combinations. In each trial, response values were generated from a spatial process with the specified parameters, and a GRTS sample and an IRS sample were selected. For the GRTS sample, the design-based approach using the local neighborhood variance (GRTS-Design) and a model-based approach were applied (GRTS-Model). For the IRS sample, the design-based approach using the simple random sample variance (IRS-Design) and a model-based approach were applied (IRS-Model).

The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package spsurvey [@dumelle2021spsurvey]. FPBK can be readily performed using the `sptotal` \textbf{\textsf{R}} package [@higham2021sptotal]. We use sptotal for both the simulation analysis and the application, estimating covariance parameters with Restricted Maximum Likelihood (REML).

We define rMS(P)E as the root-mean-squared error (design-based) or the root-mean-squared-prediction error (model-based).Figure \ref{fig:figeff} shows the relative rMS(P)E of the four approaches from Table \ref{tab:designanalysis} using the random location layout with "IRS-Design" is the baseline. More formally, the relative rMS(P)E is defined as
\begin{equation*}
\frac{\text{rMS(P)E of sampling-analysis combination}}{\text{rMS(P)E of IRS-Design}},
\end{equation*}

When there is no spatial correlation (Figure \ref{fig:figeff}, top row), the four approaches have approximately equal rMSPE, even when the assumptions of the model-based approaches are violated. So, using GRTS or using a spatial model does not result in much, if any, loss in efficiency even if the response variable is not spatially correlated. When there is high spatial correlation (Figure \ref{fig:figeff}, bottom row), the GRTS-Model approach tends to perform best, but difference in relative efficiency between GRTS-Model and GRTS-Design is small. In the lognormal, high partial sill settings (Figure \ref{fig:figeff}, bottom row and right column), IRS-Model outperforms IRS-Design by a large margin, suggesting that the poor design properties of IRS are largely mitigated by the model-based analysis.

Unsurprisingly, Figure \ref{fig:figeff} also shows that, when the assumptions for GRTS-Model are satisfied, the approach outperforms GRTS-Design. However, even when the model that generates the data is different than the model used to fit the data, as in the lognormal response, the model-based approach still outperforms the design-based approach when there is a high amount of spatial correlation. Additionally, as the sample size increases, IRS-Design performs relatively worse compared to the other approaches. These conclusions were similar to those observed when the data were gridded.

@@

```{r, fig.keep = "none"}
combo_data_grts <- combo_data %>%
  dplyr::filter(approach %in% c("Design GRTS", "Model GRTS")) %>%
  mutate(designgrts = if_else(approach == "Design GRTS", 
                              true = rmspe,
                              false = NA_real_)) %>%
  group_by(sim) %>%
  fill(designgrts, .direction = "updown") %>%
  ungroup() %>%
  mutate(releffgrts = rmspe / designgrts)
combo_data_grts_model <- combo_data_grts %>%
  filter(approach == "Model GRTS")

ggplot(data = combo_data_grts_model, aes(x = n, y = releffgrts)) +
  facet_grid(psill ~ resptype) +
  geom_point()
```

We also studied 95% interval coverage among the approaches. The design-based 95% confidence intervals and model-based 95% prediction intervals were constructed using the normal distribution. Justification for the design-based and model-based intervals comes from the asymptotic normality of totals via the Central Limit Theorem (under some assumptions).

Figure \ref{fig:figconf} shows the 95% interval coverage for each of the four approaches in the random location layout. All four approaches have somewhat similar interval coverage in all settings, with GRTS-Design having slightly lower coverage when the response is normal. 
Coverage in the normal response settings tended to be near 95%, slightly higher than coverage in the lognormal settings, which still generally exceeded 90%. Coverage for all approaches increased with the sample size for both the normal and lognormal scenarios. At a sample size of 200, all four sampling-design combinations had approximately 95% interval coverage in both response scenarios.  These conclusions were similar to those observed when the data were gridded.

```{r figconf, out.width = "100%", fig.cap = "Interval coverage for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type. The solid, black line in each plot represents 95% coverage.", message = FALSE, warning = FALSE}
ggplot(data = combo_data, aes(x = n_factor, y = coverage, colour = approach)) +
  geom_hline(yintercept = 0.95) +
  geom_jitter(width = 0.24, size = 2.5) + 
  facet_grid(psill_ratio ~ resptype) +
  scale_colour_manual(values = colour_scale) +
  ylim(0.80, 1) +
  theme_bw(base_size = 14) +
  labs(x = "Sample Size", colour = "Approach",
       y = "Interval Coverage")
```

In addition to rMS(P)E and interval coverage, we also recorded average bias. The average bias is nearly zero for all approaches in all scenarios, so we omit a visualization of the results here. The supplementary material contains tables with mean bias, rMS(P)E, and interval coverage for all 36 simulation scenarios.

>>>>

Performance of the four sampling-analysis combinations was evaluated in 36 different simulation scenarios. The 36 scenarios resulted from the crossing of three sample sizes, two location layouts, two response types, and three proportions of dependent random error. The three sample sizes ($n$) were $n = 50, n = 100,$ and $n = 200$. Samples were always selected from a population size ($N$) of $N = 900$. The two location layouts were random and gridded. Locations in the random layout were selected randomly from the unit square ($[0, 1] \times [0, 1]$). Locations in the gridded layout were selected randomly on a fixed grid from the unit square. The two response types were normal and lognormal. For the normal response type, the response was simulated using mean-zero random errors with the exponential covariance (Equation$~$\ref{equation:expcov}) for varying proportions of dependent random error. The proportion of dependent random error is represented by $\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)$, where $\sigma^2_1$ and $\sigma^2_2$ are from Equation$~$\ref{equation:expcov}. The total variance, $\sigma^2_1 + \sigma^2_2$, was always 2. The range was always $\sqrt{2} / 3$, which means that the correlation in the dependent random error decayed to nearly zero at the largest possible distance between two units in the domain. For the lognormal response type, the response was first simulated using the same approach as for the normal response type, except that the total variance was 0.6931 instead of 2. The response was then exponentiated, yielding a random variable whose total variance is 2. The lognormal responses were used to evaluate performance of the sampling-analysis approaches for data that were skewed.

```{r, echo = FALSE, results = "asis"}
n <- c("50", "100", "200")
site_locations <- c("Random", "Gridded", "")
psill_ratio <- c("0", "0.5", "0.9")
resp_type <- c("Normal", "Lognormal", "")
parm_tab <- rbind(n, site_locations, psill_ratio, resp_type)
parm_tab[c(2, 4), 3] <- "-"
row.names(parm_tab) <- c("Sample Size (n)", "Location Layout", 
                         "Proportion of Dependent Error", "Response Type")
# parm_tab <- parm_tab[c(1, 2, 4, 3), ]
parm_xtab <- xtable(parm_tab, caption = "\\label{tab:parmtab} Simulation scenario options. All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios. In each simualtion scenario, the total variance was two.")
align(parm_xtab) <- "r|lll"
print(parm_xtab, include.colnames = FALSE, hline.after = c(0, nrow(parm_xtab)), comment = FALSE)
```

In each of the 36 simulation scenarios, there were 2000 independent simulation trials. In each trial, IRS and GRTS samples were selected and then design-based and model-based analyses were used to estimate the mean and construct confidence (design-based) or prediction (model-based) intervals. We recorded the bias, squared error, and interval coverage for all sampling-analysis combinations in each trial. Then we summarized the performance of the combinations across trials by calculating average bias, rMS(P)E (root-mean-squared error for the design-based approaches and root-mean-squared-prediction error for the model-based approaches), and the rate at which the true mean is contained in its 95% interval. The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package `spsurvey` [@dumelle2021spsurvey]. FPBK is available in the `sptotal` \textbf{\textsf{R}} package [@higham2021sptotal] and covariance parameters were estimated using Restricted Maximum Likelihood  [@patterson1971recovery; @harville1977maximum; @wolfinger1994computing].

Figure \ref{fig:figeff} shows the relative rMS(P)E of the four approaches from Table \ref{tab:designanalysis} using the random location layout with "IRS-Design" is the baseline. More formally, the relative rMS(P)E is defined as
\begin{equation*}
\frac{\text{rMS(P)E of sampling-analysis combination}}{\text{rMS(P)E of IRS-Design}},
\end{equation*}
When there is no spatial correlation (Figure \ref{fig:figeff}, top row), the four sampling-analysis combinations have approximately equal rMS(P)E. So, using GRTS or using a spatial model does not result in much, if any, loss in efficiency even when the response variable is not spatially correlated. When there is spatial correlation (Figure \ref{fig:figeff}, middle and bottom row), the GRTS-Model combination tends to perform best, followed by GRTS-Design, IRS-Model, and finally IRS-Design, though the difference in relative rMS(P)E among IRS-Model, GRTS-Design, and GRTS-Model is relatively small.  As the strength of spatial correlation increases, the gap in rMS(P)E between IRS-Design and the other combinations widens. Finally we note that when there is spatial correlation, IRS-Model outperforms IRS-Design by a large margin, suggesting that the poor design properties of IRS are largely mitigated by the model-based analysis. These conclusions are similar to those observed in the grid location layout. Tables for rMS(P)E in all 36 simulation scenarios are provided in the supplementary material.

@@

We also studied 95% interval coverage among the combinations. The design-based 95% confidence intervals and model-based 95% prediction intervals were constructed using the normal distribution. Justification for the design-based and model-based intervals comes from the asymptotic normality of totals via the Central Limit Theorem.

Figure \ref{fig:figconf} shows the 95% interval coverage for each of the four combinations in the random location layout. All four combinations have fairly similar interval coverage within each scenario. Coverage in the normal response scenarios tended to be near 95% and slightly higher than coverage in the lognormal scenarios. Coverage in the lognormal scenarios still generally exceeded 90%. Coverage tended to always increase with the sample size. At a sample size of 200, all four combinations had approximately 95% interval coverage in both response scenarios and all dependent error proportions.  These conclusions were similar to those found in the grid location layout.
Tables for interval coverage in all 36 simulation scenarios are provided in the supplementary material.

```{r figconf, out.width = "100%", fig.cap = "Interval coverage for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type. The solid, black line in each plot represents 95\\% coverage.", message = FALSE, warning = FALSE}
ggplot(data = combo_data, aes(x = n_factor, y = coverage, colour = approach)) +
  geom_hline(yintercept = 0.95) +
  geom_jitter(width = 0.24, size = 2.5) + 
  facet_grid(psill_ratio ~ resptype) +
  scale_colour_manual(values = colour_scale) +
  ylim(0.80, 1) +
  theme_bw(base_size = 14) +
  labs(x = "Sample Size", colour = "Approach",
       y = "Interval Coverage")
```

The average bias was nearly zero for all four combinations in all 36 scenarios, so we omit a visualization of those results here. Tables for average bias in all 36 simulation scenarios are provided in the supplementary material.

### 2021-11-12 inst/manuscript/manuscript.tex
section 3 updates

Performance of the four sampling-analysis combinations were evaluated in
36 different simulation scenarios. The 36 scenarios resulted from the
crossing of three sample sizes, two location layouts, two response
types, and three proportions of dependent random error. The three sample
sizes (\(n\)) were \(n = 50, n = 100,\) and \(n = 200\). Samples were
always selected from a population size (\(N\)) of \(N = 900\). The two
location layouts were random and gridded. Locations in the random layout
were selected randomly from the unit square (\([0, 1] \times [0, 1]\)).
Locations in the gridded layout were selected randomly on a fixed grid
from the unit square. The two response types were normal and lognormal.
For the normal response type, the response was simulated using mean-zero
random errors with the exponential covariance
(Equation\(~\)\ref{equation:expcov}) for varying proportions of
dependent random error. The proportion of dependent random error is
represented by \(\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)\), where
\(\sigma^2_1\) and \(\sigma^2_2\) are from
Equation\(~\)\ref{equation:expcov}. The total variance
(\(\sigma^2_1 + \sigma^2_2\)) was always 2. The range was always
\(\sqrt{2} / 3\), which means that the correlation in the dependent
random error decays to nearly zero at the largest possible distance
between two units in the domain. For the lognormal response type, the
response was first simulated using the same approach as for the normal
response type, except that the total variance was 0.6931 instead of 2.
The response was then exponentiated, yielding a total variance of 2 on
the exponentiated scale. The lognormal responses were used to evaluate
performance of the sampling-analysis approaches for data that were
skewed.

a crossed design with the simulation parameters given in Table
\ref{tab:parmtab} for a total of 36 scenarios. All scenarios used
exponential correlation with a \(\sqrt{4} / 3\) for \(N = 900\) response
values simulated on the unit square in either random locations (Layout =
Random) or gridded locations (Layout = Gridded). The mean for the
spatial process generating the response was set to zero.

For the lognormal scenarios, the response values were simulated using
the specified correlation parameters using a normal distribution and
were subsequently exponentiated. A total variance of 0.6931 and a mean
of 0 on the normal scale yielded a total variance of 2 and a mean of
1.414 after exponentiation. Therefore, when the model-based methods were
used for lognormal response, the correlation was mis-specified. We chose
to simulate values with a lognormal distribution so that we could test
the model-based analysis approach with a mis-specified model and so that
we could test both analysis approaches on data that exhibits a large
amount of skewness.

\begin{table}[ht]
\centering
\begin{tabular}{r|lll}
   \hline
Sample Size (n) & 50 & 100 & 200 \\ 
  Location Layout & Random & Gridded & - \\ 
  Proportion of Dependent Error & 0 & 0.5 & 0.9 \\ 
  Response Type & Normal & Lognormal & - \\ 
   \hline
\end{tabular}
\caption{\label{tab:parmtab} Simulation scenario options. All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios. In each simualtion scenario, the mean was zero and the variance was two.} 
\end{table}

There were 2000 simulation trials for each of the 36 parameter
combinations. In each trial, response values were generated from a
spatial process with the specified parameters, and a GRTS sample and an
IRS sample were selected. For the GRTS sample, the design-based approach
using the local neighborhood variance (GRTS-Design) and a model-based
approach were applied (GRTS-Model). For the IRS sample, the design-based
approach using the simple random sample variance (IRS-Design) and a
model-based approach were applied (IRS-Model).

The GRTS algorithm and the local neighborhood variance estimator are
available in the \textbf{\textsf{R}} package spsurvey (Dumelle et al.,
2021). FPBK can be readily performed using the \texttt{sptotal}
\textbf{\textsf{R}} package (Higham et al., 2021). We use sptotal for
both the simulation analysis and the application, estimating covariance
parameters with Restricted Maximum Likelihood (REML).

We define rMS(P)E as the root-mean-squared error (design-based) or the
root-mean-squared-prediction error (model-based).Figure \ref{fig:figeff}
shows the relative rMS(P)E of the four approaches from Table
\ref{tab:designanalysis} using the random location layout with
``IRS-Design'' is the baseline. More formally, the relative rMS(P)E is
defined as \begin{equation*}
\frac{\text{rMS(P)E of sampling-analysis combination}}{\text{rMS(P)E of IRS-Design}},
\end{equation*}

When there is no spatial correlation (Figure \ref{fig:figeff}, top row),
the four approaches have approximately equal rMSPE, even when the
assumptions of the model-based approaches are violated. So, using GRTS
or using a spatial model does not result in much, if any, loss in
efficiency even if the response variable is not spatially correlated.
When there is high spatial correlation (Figure \ref{fig:figeff}, bottom
row), the GRTS-Model approach tends to perform best, but difference in
relative efficiency between GRTS-Model and GRTS-Design is small. In the
lognormal, high partial sill settings (Figure \ref{fig:figeff}, bottom
row and right column), IRS-Model outperforms IRS-Design by a large
margin, suggesting that the poor design properties of IRS are largely
mitigated by the model-based analysis.

Unsurprisingly, Figure \ref{fig:figeff} also shows that, when the
assumptions for GRTS-Model are satisfied, the approach outperforms
GRTS-Design. However, even when the model that generates the data is
different than the model used to fit the data, as in the lognormal
response, the model-based approach still outperforms the design-based
approach when there is a high amount of spatial correlation.
Additionally, as the sample size increases, IRS-Design performs
relatively worse compared to the other approaches. These conclusions
were similar to those observed when the data were gridded.

@@

We also studied 95\% interval coverage among the approaches. The
design-based 95\% confidence intervals and model-based 95\% prediction
intervals were constructed using the normal distribution. Justification
for the design-based and model-based intervals comes from the asymptotic
normality of totals via the Central Limit Theorem (under some
assumptions).

Figure \ref{fig:figconf} shows the 95\% interval coverage for each of
the four approaches in the random location layout. All four approaches
have somewhat similar interval coverage in all settings, with
GRTS-Design having slightly lower coverage when the response is normal.
Coverage in the normal response settings tended to be near 95\%,
slightly higher than coverage in the lognormal settings, which still
generally exceeded 90\%. Coverage for all approaches increased with the
sample size for both the normal and lognormal scenarios. At a sample
size of 200, all four sampling-design combinations had approximately
95\% interval coverage in both response scenarios. These conclusions
were similar to those observed when the data were gridded.

\textbackslash begin\{figure\}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figconf-1}
\textbackslash caption\{Interval coverage for the four sampling-analysis
combinations. The rows indicate the proportion of dependent error and
the columns indicate the response type. The solid, black line in each
plot represents 95\% coverage.\}\label{fig:figconf}
\textbackslash end\{figure\}

In addition to rMS(P)E and interval coverage, we also recorded average
bias. The average bias is nearly zero for all approaches in all
scenarios, so we omit a visualization of the results here. The
supplementary material contains tables with mean bias, rMS(P)E, and
interval coverage for all 36 simulation scenarios.

@@

@@

@@

>>>>

Performance of the four sampling-analysis combinations was evaluated in
36 different simulation scenarios. The 36 scenarios resulted from the
crossing of three sample sizes, two location layouts, two response
types, and three proportions of dependent random error. The three sample
sizes (\(n\)) were \(n = 50, n = 100,\) and \(n = 200\). Samples were
always selected from a population size (\(N\)) of \(N = 900\). The two
location layouts were random and gridded. Locations in the random layout
were selected randomly from the unit square (\([0, 1] \times [0, 1]\)).
Locations in the gridded layout were selected randomly on a fixed grid
from the unit square. The two response types were normal and lognormal.
For the normal response type, the response was simulated using mean-zero
random errors with the exponential covariance
(Equation\(~\)\ref{equation:expcov}) for varying proportions of
dependent random error. The proportion of dependent random error is
represented by \(\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)\), where
\(\sigma^2_1\) and \(\sigma^2_2\) are from
Equation\(~\)\ref{equation:expcov}. The total variance,
\(\sigma^2_1 + \sigma^2_2\), was always 2. The range was always
\(\sqrt{2} / 3\), which means that the correlation in the dependent
random error decayed to nearly zero at the largest possible distance
between two units in the domain. For the lognormal response type, the
response was first simulated using the same approach as for the normal
response type, except that the total variance was 0.6931 instead of 2.
The response was then exponentiated, yielding a random variable whose
total variance is 2. The lognormal responses were used to evaluate
performance of the sampling-analysis approaches for data that were
skewed.

\begin{table}[ht]
\centering
\begin{tabular}{r|lll}
   \hline
Sample Size (n) & 50 & 100 & 200 \\ 
  Location Layout & Random & Gridded & - \\ 
  Proportion of Dependent Error & 0 & 0.5 & 0.9 \\ 
  Response Type & Normal & Lognormal & - \\ 
   \hline
\end{tabular}
\caption{\label{tab:parmtab} Simulation scenario options. All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios. In each simualtion scenario, the total variance was two.} 
\end{table}

In each of the 36 simulation scenarios, there were 2000 independent
simulation trials. In each trial, IRS and GRTS samples were selected and
then design-based and model-based analyses were used to estimate the
mean and construct confidence (design-based) or prediction (model-based)
intervals. We recorded the bias, squared error, and interval coverage
for all sampling-analysis combinations in each trial. Then we summarized
the performance of the combinations across trials by calculating average
bias, rMS(P)E (root-mean-squared error for the design-based approaches
and root-mean-squared-prediction error for the model-based approaches),
and the rate at which the true mean is contained in its 95\% interval.
The GRTS algorithm and the local neighborhood variance estimator are
available in the \textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle
et al., 2021). FPBK is available in the \texttt{sptotal}
\textbf{\textsf{R}} package (Higham et al., 2021) and covariance
parameters were estimated using Restricted Maximum Likelihood (Harville,
1977; Patterson and Thompson, 1971; Wolfinger et al., 1994).

Figure \ref{fig:figeff} shows the relative rMS(P)E of the four
approaches from Table \ref{tab:designanalysis} using the random location
layout with ``IRS-Design'' is the baseline. More formally, the relative
rMS(P)E is defined as \begin{equation*}
\frac{\text{rMS(P)E of sampling-analysis combination}}{\text{rMS(P)E of IRS-Design}},
\end{equation*} When there is no spatial correlation (Figure
\ref{fig:figeff}, top row), the four sampling-analysis combinations have
approximately equal rMS(P)E. So, using GRTS or using a spatial model
does not result in much, if any, loss in efficiency even when the
response variable is not spatially correlated. When there is spatial
correlation (Figure \ref{fig:figeff}, middle and bottom row), the
GRTS-Model combination tends to perform best, followed by GRTS-Design,
IRS-Model, and finally IRS-Design, though the difference in relative
rMS(P)E among IRS-Model, GRTS-Design, and GRTS-Model is relatively
small. As the strength of spatial correlation increases, the gap in
rMS(P)E between IRS-Design and the other combinations widens. Finally we
note that when there is spatial correlation, IRS-Model outperforms
IRS-Design by a large margin, suggesting that the poor design properties
of IRS are largely mitigated by the model-based analysis. These
conclusions are similar to those observed in the grid location layout.
Tables for rMS(P)E in all 36 simulation scenarios are provided in the
supplementary material.

@@

We also studied 95\% interval coverage among the combinations. The
design-based 95\% confidence intervals and model-based 95\% prediction
intervals were constructed using the normal distribution. Justification
for the design-based and model-based intervals comes from the asymptotic
normality of totals via the Central Limit Theorem.

Figure \ref{fig:figconf} shows the 95\% interval coverage for each of
the four combinations in the random location layout. All four
combinations have fairly similar interval coverage within each scenario.
Coverage in the normal response scenarios tended to be near 95\% and
slightly higher than coverage in the lognormal scenarios. Coverage in
the lognormal scenarios still generally exceeded 90\%. Coverage tended
to always increase with the sample size. At a sample size of 200, all
four combinations had approximately 95\% interval coverage in both
response scenarios and all dependent error proportions. These
conclusions were similar to those found in the grid location layout.
Tables for interval coverage in all 36 simulation scenarios are provided
in the supplementary material.

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figconf-1} \caption{Interval coverage for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type. The solid, black line in each plot represents 95\% coverage.}\label{fig:figconf}
\end{figure}

The average bias was nearly zero for all four combinations in all 36
scenarios, so we omit a visualization of those results here. Tables for
average bias in all 36 simulation scenarios are provided in the
supplementary material.

@@

\leavevmode\hypertarget{ref-harville1977maximum}{}%
Harville, D.A., 1977. Maximum likelihood approaches to variance
component estimation and to related problems. Journal of the American
statistical association 72, 320--338.

@@

\leavevmode\hypertarget{ref-patterson1971recovery}{}%
Patterson, H.D., Thompson, R., 1971. Recovery of inter-block information
when block sizes are unequal. Biometrika 58, 545--554.

@@

\leavevmode\hypertarget{ref-wolfinger1994computing}{}%
Wolfinger, R., Tobias, R., Sall, J., 1994. Computing gaussian
likelihoods and their derivatives for general linear mixed models. SIAM
Journal on Scientific Computing 15, 1294--1310.
